---
title: "Survey II Assignment"
output: html_document
date: "2025-02-13"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r}
library(tidyverse)
library(tidyr)
library(haven)
library(readxl)
library(xml2)
library(rvest)
library(janitor)
library(DataExplorer)
library(countrycode)

```

## Loading data

### Survey data

```{r}
data_raw <- read_dta("ZA7575.dta")
```

### GDP per capita

Retrieved from Eurostat for the year 2019
(<https://ec.europa.eu/eurostat/databrowser/view/tipsna40/default/table?lang=en>).

```{r}
gdp_data <- read_xlsx("gdp_pc_2019.xlsx", sheet = 3, range = "A11:B39", col_names = c("country", "gdp_pc"))
head(gdp_data)
```

----- United Kingdom is missing from this dataset as well as from a few
others in Eurostat (we might have to find it somewhere else).

### Rural population

Retrieved from the World Bank
(<https://data.worldbank.org/indicator/SP.RUR.TOTL.ZS>). Data on the %
of population living in rural areas.

```{r}
rural_data <- read_xls("rural_pop.xls", sheet = 1, range = "A4:BL270")

rural_data <- rural_data |> 
  select(`Country Name`, `Country Code`, `2019`) |> 
  clean_names() |> 
  rename("rural_pop_percentage" = "x2019")

# select European countries
codelist <- countrycode::codelist |> 
  select(country.name.en, cow.name, eu28, iso.name.en, iso2c, iso3c, un.name.en) |> # selecting variables that might be useful for joins (will probably have to clean later on)
  filter(!is.na(eu28)) # including the countries that belonged to the EU in 2019 which are listed under Q1 of the questionnaire

rural_data <- rural_data |> 
  inner_join(select(codelist, iso3c), by = c("country_code" = "iso3c"))
```

### Same sex unions

Retrieved from Wikipedia.

```{r}
link <- "https://en.wikipedia.org/wiki/Recognition_of_same-sex_unions_in_Europe"
html_website <- link |> read_html()
all_tables <- html_website|> html_table()
marriage_data <- all_tables[[4]]
marriage_data <- marriage_data |> select(Status, Country)

# cleaning the status column
marriage_data$Status <- sub("\\s*\\(.*|\\s*-.*", "", marriage_data$Status)
unique(marriage_data$Status)

# cleaning the country column
marriage_data <- marriage_data |> 
  mutate(Country = str_replace_all(Country, c("\\*" = "", "†" = "", "\\[.*?\\]" = ""))) |> 
  mutate(Country = trimws(Country))
unique(marriage_data$Country)

# getting rid of the total and subtotal rows
marriage_data <- marriage_data |> 
  filter(!Status %in% c("Total", "Subtotal"))

# again keeping only the countries that appear in the questionnaire
marriage_data <- marriage_data |> 
  inner_join(select(codelist, cow.name), by = c("Country" = "cow.name"))
```

This variable could be recoded as either ordinal, categorical (both
factor) or logical.

------- The dataset has 32 observations instead of 28 (EU-28 countries).
According to this test, Croatia, Hungary, Latvia and Slovakia have two
different types of status, so maybe we should cross-check with other
sources?

```{r}
marriage_data |> 
  add_count(Country) |> 
  filter(n > 1)
```

### Gender Equality Index

Developed by the European Institute for Gender Equality. Retrieved from
<https://eige.europa.eu/gender-statistics/dgs/indicator/index_data__index_scores/datatable>
for the year 2019.

```{r}
gequality_index <- read_xlsx("gender_equality_index_scores.xlsx", range = "A16:V44")
names(gequality_index) <- names(gequality_index) %>%
 janitor::make_clean_names() %>%
  gsub("_score$", "", .)
names(gequality_index)

gequality_index <- gequality_index |>  
  rename(country_name = geographic_region_sub_domain_scores)
```

###Democracy Index Using the Economist's Democracy Index (there might be
a better one?).

```{r}
democracy_index <- read_xlsx("EIU_democracy_index.xlsx", sheet = 4)

# the ISO codes were lowercase which impedes the join
democracy_index$geo <- toupper(democracy_index$geo)

# filter for 2019 and EU28 countries
democracy_index <- democracy_index |> 
  filter(time == 2019) |> 
  inner_join((select(codelist, iso3c)), by = c("geo" = "iso3c"))

# clean var names
names(democracy_index) <- names(democracy_index) %>%
 janitor::make_clean_names() %>%
  gsub("_eiu$", "", .)

names(democracy_index)

democracy_index <- democracy_index |> 
  rename(country_code = geo,
         country_name = name,
         year = time)
```

In both of these tables (Gender Equality and Democracy indexes) there
are more indicators apart from the overall indexes, we have to decide if
we think they're useful or if we can discard some/all of them.

## Initial data cleaning

### Check overall data quality

tbc for initial checks?

### Country data

We have columns for West and East Germany in our data. To match the
iso2c classifications, we join these in the raw data. may want to
consolidate all of the groups. There is a separate column called "United
Germany" that is a dummy variable.

Easiest way to fix this is just only take the first 2 letters of the
country codes. We then get the EU28 countries.

```{r}
# extract labels - we see 40 countries but it specifies which are not includes.
country_labels <- attr(data_raw$country, "labels")
# join on names
country_tibble <- tibble(cntry_code = as.integer(country_labels), 
                         country_name = names(country_labels))
country_tibble

# clean country code ID and then join on names from codelist table
data_raw <- data_raw |> 
  mutate(countryid = str_sub(isocntry, 1, 2),
         country_names = countrycode(countryid, 
                                               origin = "iso2c", 
                                               destination = "country.name"), 
         .after=isocntry) 

# #Now check the summary of our countries. Which matches the EU28,
data_raw |> 
  group_by(countryid, country_names) |> 
  tally() #count the observations too

```

# Identify key socio-demogrpahic variables

From exploration of the codebook and dataframe:

*Target variable:* \* qc19 "Do you think that transgender persons should
be able to change their civil documents to match their inner gender
identity?"

Survey ID variables: - studyno1 to survey vars, all to be removed.

Respondent information -\> questions starting with "D" (D1-D77) -
political views, living situation, gender, education, occupation,
household characteristics, class

Key socioeconomic variables:

-   d1 - political ideology, left to right

-   d10 - Gender

-   d11 - Age exact

    -   Alternatives: d11r1 age recoded 4 cats, d11r2 age recoded 6
        cats, d11r3 age recoded 7 cats

-   d8 - Education, how old were you when you stopped full time
    education (Student = 00; No education = 01; refusal = 98; DK = 99

    -   d8r1 -\> 11 category recode

    -   d8r2 -\> 5 category recode s

-   d3 - religion

-   d25 - region lived in (1=rural, 2=small/medium town, 3=large town,
    4=DK)

-   d70 - life satisfaction: - 1 = very satisfied, 2= fairly, 3=not
    very, 4 = not at all, 5 = DK

-   D63 - class (working to higher, 1;5 then 6-9 no resp)

Uncertain variables, marital status (d7) /hh arrangement (d40)
/occupation (d15) / internet use (d60) Also the "sd" prefix variables..
these are about contact with different groups of discriminated people?

Paradata information -\> questions starting with "P" Paradata: -\>
P1-P10, date, time, length of interview, people present during
interview, respondent cooperation, location vars, interviewer number.

Select and rename these key variables:

```{r}
individuals_df <- data_raw |> 
  
  #use select function and rename all the in scope variables 
  select(
      uniqid, 
      isocntry,
      countryid,
      country_names,
        
    
      target_var = qc19,
      
    # socioeconomic variables for individuals 
      gender = d10,
      age = d11,
      age_4cat = d11r1,
      age_6cat = d11r2,
      age_7cat = d11r3,
      polit_ideology = d1,
      educ_5cat = d8r2, # all the education vars have low cutoffs 
      educ_11cat = d8r1,
      lifesat = d70,
      religion = sd3, 
      class = d63,
      occ_group_recode = d15a_r2, # the occupation vars are maybe not needed too
      financialstress = d60, #difficulty paying bills
      polintr, # no need to rename political interest var
      
    # paradata variables   
      int_date = p1,
      int_time = p2,
      int_length = p3r, #using recoded var instead of base p3 
      int_ppl_present = p4,
      int_resp_coop = p5,
      int_location = p6, # this may be a repeat of locality variable in survey. Useful maybe to consolidate data
      #p7 not useful, location, no data for p8-11
      
      #p13 is language of interview. Probably not useful? 
      )

```

Misc code for checking the sociodemographic vars (to be deleted)

```{r}
#check repeated cols
data_raw |> 
  select(starts_with("p13"))

# check labels
attr(data_raw$d15a_r2, "labels")
attr(data_raw$d7r, "labels")

#check values
summary(data_raw$p9)

```

## Remove unwanted variables

The approach above may be best.. but I did start this first by deleting
variables by process of elimination. Maybe good to justify some
decisions, but I'm not sure it matters either way.

-   we do need to decide on some of the attitudinal questions if they
    are relevant enough to include. I haven't looked at all enough in
    detail.

Variable groups I suggest removing: - gen1:gen6 -\> generation
variables, offer no more info than age - nationality variables -\> would
be easier and more accurate to join an overall % of immigration variable
if we want this data. - d71, -\> the three variables are summarised in
the `polintr` variable.

```{r}
id_vars <- c("studyno1", "studyno2", "doi", "version", "edition", "survey", "caseid", "uniqid", "serialid", "tnscntry", "country")
# remove ID vars
clean1 <- data_raw |> 
  select(-any_of(id_vars))

#remove nationality variable (q1). Later, if we decide we need an immigration variable, we can impute national level data.  
clean1 <- clean1 |> 
  select(-any_of(starts_with("q1")))

# remove generation variable, too similar to age
clean1 <- clean1 |> 
  select(-any_of(starts_with("gen")))

# remove additional paradata
clean1 <- clean1 |> 
  select(-any_of(starts_with("p7", "p13")))

```

To discuss re: cleaning. But I think also remove all these - qa set
(from qa1 to qa20) \* on trade, globalisation, eu business

-   qb set (from qb1 to qb9) \* on energy policy and EU priorities

Do we delete all these or are we expected to do some sort of
unsupervised learning or similar to consider their attitudes? I think
not.. There is questions about priority for business

Checking how many more variables it removes. It would be 159 removed.

```{r}
data_raw |> 
  select(any_of(starts_with(c("qa", "qb")))) |> 
  ncol()
```

Removing from my test data, this still has 409 variables though.

```{r}
clean1 <- clean1 |> 
  select(-any_of(starts_with(c("qa", "qb"))))

```

## Clean respondents personal data

I think some recoding or factorising may be necessary.

Note: There are lots of haven labelled columns.. it would be quite easy
to write a function to extract the label names to a new column? This
could be easier for when we want to investigate new variables?

```{r}

```

### Impute actual NA values 

As these are inconsistent between each variable, I guess we need to
override and decide if we then impute the NA values.

Clean the education variable. Currently we have: d8 - Education, how old
were you when you stopped full time education (Student = 00; No
education = 01; refusal = 98; DK = 99 - d8r1 -\> 11 category recode -
d8r2 -\> 5 category recode

Clean the

```{r}
table(data_raw$d8, data_raw$d8r2)

# First, recode the d8 values that are given as NA or student/no education
clean1 |> 
  mutate(educ = if_else(d8 %in% c(0, 1, 98, 99), NA, d8))

# if d8 = 0, they are a current student. 

```

### Analysis of missing values.

Marga discussed the importance of understanding missing values. I think
this will be an important section where we also model what type of
people are more/less likely to not respond. It's a sort of robustness
check to see how much we can trust the models. Or if specific
demographics of people are less likely to respond, then we may trust it
less.

E.g. if men/religious people are less likely to respond, it could be an
issue. Ideally we would test the representativeness of our respondents
to the questions about transgender discrimination and the overall
populations.

```{r}
plot_missing()
summary(data_raw$d11)
# returns nothing now because of NA values are all hard coded as mixed values e.g. some are 7, 98, 12. We probably need to standardise, section above maybe
  
```

## Descriptive data analysis and checks to start

Run comparisons by country to compare ages, gender distribution, Number
of respondents per country

```{r}
# number of respondents 
```

## Testing replication of the initial plot:

-   basic version only.

```{r}

```

-   we could also recreate without any NA values to see if it makes the
    plot just seem visually more even to start?

## Modelling

Some initial thoughts:

a)  modelling the target variables
    -   base model -\> using linear mixed model to analyse individual
        and country level differences

    -   model including paradata -\> see if these impacted outcomes too

b\. modelling differences in non-response?

-   I feel like Marga mentioned this? we would need to just create a
    dummy variable for NA response to the target variable question. This
    is our secondary model target variable. I guess this could be a
    classification model? for discussion but also more thoughts in my
    missing data note above...
