---
title: "Survey II Assignment"
output: html_document
date: "2025-02-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries
```{r}
library(tidyverse)
library(tidyr)
library(haven)
library(readxl)
library(xml2)
library(rvest)
library(janitor)
```

## Loading data

### Survey data
```{r}
data_raw <- read_dta("ZA7575.dta")
```

### GDP per capita
Retrieved from Eurostat for the year 2019 (https://ec.europa.eu/eurostat/databrowser/view/tipsna40/default/table?lang=en).
```{r}
gdp_data <- read_xlsx("gdp_pc_2019.xlsx", sheet = 3, range = "A11:B39", col_names = c("country", "gdp_pc"))
head(gdp_data)
```
----- United Kingdom is missing from this dataset as well as from a few others in Eurostat (we might have to find it somewhere else).

### Rural population
Retrieved from the World Bank (https://data.worldbank.org/indicator/SP.RUR.TOTL.ZS). Data on the % of population living in rural areas.
```{r}
rural_data <- read_xls("rural_pop.xls", sheet = 1, range = "A4:BL270")

rural_data <- rural_data |> 
  select(`Country Name`, `Country Code`, `2019`) |> 
  clean_names() |> 
  rename("rural_pop_percentage" = "x2019")

# select European countries
codelist <- countrycode::codelist |> 
  select(country.name.en, cow.name, eu28, iso.name.en, iso3c, un.name.en) |> # selecting variables that might be useful for joins (will probably have to clean later on)
  filter(!is.na(eu28)) # including the countries that belonged to the EU in 2019 which are listed under Q1 of the questionnaire

rural_data <- rural_data |> 
  inner_join(select(codelist, iso3c), by = c("country_code" = "iso3c"))
```

### Same sex unions
Retrieved from Wikipedia.
```{r}
link <- "https://en.wikipedia.org/wiki/Recognition_of_same-sex_unions_in_Europe"
html_website <- link |> read_html()
all_tables <- html_website|> html_table()
marriage_data <- all_tables[[4]]
marriage_data <- marriage_data |> select(Status, Country)

# cleaning the status column
marriage_data$Status <- sub("\\s*\\(.*|\\s*-.*", "", marriage_data$Status)
unique(marriage_data$Status)

# cleaning the country column
marriage_data <- marriage_data |> 
  mutate(Country = str_replace_all(Country, c("\\*" = "", "†" = "", "\\[.*?\\]" = ""))) |> 
  mutate(Country = trimws(Country))
unique(marriage_data$Country)

# getting rid of the total and subtotal rows
marriage_data <- marriage_data |> 
  filter(!Status %in% c("Total", "Subtotal"))

# again keeping only the countries that appear in the questionnaire
marriage_data <- marriage_data |> 
  inner_join(select(codelist, cow.name), by = c("Country" = "cow.name"))
```
This variable could be recoded as either ordinal, categorical (both factor) or logical.


------- The dataset has 32 observations instead of 28 (EU-28 countries). According to this test, Croatia, Hungary, Latvia and Slovakia have two different types of status, so maybe we should cross-check with other sources?
```{r}
marriage_data |> 
  add_count(Country) |> 
  filter(n > 1)
```

### Gender Equality Index
Developed by the European Institute for Gender Equality. Retrieved from https://eige.europa.eu/gender-statistics/dgs/indicator/index_data__index_scores/datatable for the year 2019.

```{r}
gequality_index <- read_xlsx("gender_equality_index_scores.xlsx", range = "A16:V44")
names(gequality_index) <- names(gequality_index) %>%
 janitor::make_clean_names() %>%
  gsub("_score$", "", .)
names(gequality_index)

gequality_index <- gequality_index |>  
  rename(country_name = geographic_region_sub_domain_scores)
```
###Democracy Index
Using the Economist's Democracy Index (there might be a better one?).

```{r}
democracy_index <- read_xlsx("EIU_democracy_index.xlsx", sheet = 4)

# the ISO codes were lowercase which impedes the join
democracy_index$geo <- toupper(democracy_index$geo)

# filter for 2019 and EU28 countries
democracy_index <- democracy_index |> 
  filter(time == 2019) |> 
  inner_join((select(codelist, iso3c)), by = c("geo" = "iso3c"))

# clean var names
names(democracy_index) <- names(democracy_index) %>%
 janitor::make_clean_names() %>%
  gsub("_eiu$", "", .)

names(democracy_index)

democracy_index <- democracy_index |> 
  rename(country_code = geo,
         country_name = name,
         year = time)
```
In both of these tables (Gender Equality and Democracy indexes) there are more indicators apart from the overall indexes, we have to decide if we think they're useful or if we can discard some/all of them.

## Data cleaning

